{\it Maxwell's Silver Hammer:}\\ On pulsed microwave acoustic resonant viral inactivation

### 2021-02-20 documents/experiment_1_data.tex
whatever
test

\begin{center}
	\begin{tabular}{ |c|c|c| } 
		\hline
		 & col2 & col3 \\
		\hline
		\multirow{3}{4em}{Multiple row} 
		& cell2 & cell3 \\ 
		& cell5 & cell6 \\ 
		& cell8 & cell9 \\ 
		\hline
	\end{tabular}
\end{center}

>>>>

\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c| } 
		\hline
		col1 & kilocounts per 10s integration \\
		\hline
		Autoclaved (0.2 mL) & 32 \\ 
		Autoclaved (50 uL, sham) & 30 \\ 
		Autoclaved (0.2 mL) & 29 \\ 
		PG1 'blank' & 8 \\ 
		Dist. water & 0.5 \\
		\hline
		\multicolumn{2}{|c|}{\ntilde 0.25 MV/m, 0.2 uL/s at 500 Hz  } \\
		\hline

		& cell5  \\ 
		& Negative  \\ 
		\hline
	\end{tabular}

\end{table}

Data from testing on bacteriophage. Data rounded to the nearest thousand photon counts per 10 seconds. Not thoroughly characterized.

Field strengths are very approximate.

Background ('blank', in notes) readings measure the amount of DNA present in the phage culture tube before any treatment.

Repeatability between test conditions was not excellent. The source of the ~7000 count variation between backgrounds is not known.

The flurometer printed a constant stream of results, all within the random error between capture windows $\sigma=0.5$ kcounts. Only after most of the data was taken was it realized how dumb this was.

### 2021-02-20 documents/summary.tex
test

We have some minor questions regarding experimental technique\footnote{The use of absorption spectroscopy methods without an error analysis is concerning (see below); an effective charge of $q=10^7 e^-$ was found, which is very hard to explain; a lack of free-space dosimetry or simulation means frequency-selective inactivation could have been due to instrument response, infrared thermography only determines surface temperature.} in the original papers that can only be addressed by more extensive replication studies, and luckily are due to be answered by other groups\cite{Generating}, so we do not defer greatly to their observations.

@@

The genome charge is almost entirely neutralized  From an energetics perspective, \footnote{we can obtain the spring constant $k = \omega_{res}^2 m^*$ from the resonant frequency and therefore from the speed of sound and $m^*$, the reduced mass.} With relatively sane choices of charge screening and resonant frequency, the energy absorbed per monocycle at 5 MV/m varies from 0.0001 to 200 times the thermal energy $k_b T$, so the maximum energy

Compare to ~5 $k_b T$ per protein for 1 capsid segment binding energy \cite{Energies2012} \cite{Weak2002} and 6 to 30 $k_b T$ to form a single pore in a lipid bilayer \cite{Atomistic2014a}.
%17 or 78 kJ/mol / (boltzmannConstant * 310 K) / avogadroConstant = 6.5 to 30.3 kT

On the other hand, a 0.5 angstrom movement is only a 0.04\% deflection of the capsid, whereas a general value for host cells is about 3\% deflection. In addition, with a relaxation time on the order of 500 ps, it is difficult to imagine that the energy will be sufficiently concentrated to be destructive, as after the relaxation time the energy is a distant memory; only the mechanical effects that were induced during that time can remain.

@@

note clinical irrev electroporation

@@

In terms of host safety, Pakhomov \cite{Comparative} find no non-thermal effects on functioning frog heart pacemaker cells at 0.9 MV/m (although the field inside the tissue was not measured and could have been lower). Excellent recent \textit{in vivo} data by de Seze following continuous exposure to nanosecond pulse trains at an external field up to 3 MV/m \cite{Repeated2020} notes little acute effects, but a massive rate of long-term carcinogenicity.

@@

It is difficult to make spark gaps run faster than 200 ps without high-pressure or Marx overvoltage. Avalanche transistors appear to be the easiest way to produce high power sub 100 picosecond pulses. We stole an avalanche Marx design verbatim from Li (\cite{Development2016b} \cite{Design2018c}), which allows large pulses from safer lower voltage supplies (albeit requiring somewhat costly capacitors); but

@@

We are particularly excited by the prospect of using picosecond time-domain measurements not only to perform steady state dielectric measurements or monitor the input pulse, as is now common practice, but to resolve the internal effects of a single high-voltage pulse on the virus in real time. Scan-converter based transient digitizers, such as the Tektronix SCD-5000 series, appear to be the lowest-cost method of obtaining such THz sampling bandwidths. When high voltage pulses are available, miniaturized \ntilde 2 mm internal diameter continuously pumped scan converter tubes like Fecher micro-oscillographs using in-vacuum CMOS image sensors for direct-electron-detection appear to be an easy way to digitize 20+ GHz transients. The main bandwidth limitation appears to be the length of the Y deflection plate and the sensitivity of the recording medium.

>>>>

We have some minor questions regarding experimental technique in the original papers that can only be addressed by more extensive replication studies, and luckily are due to be answered by other groups\cite{Generating}, so we do not defer greatly to their observations.

@@

Because the effect is so remarkable, we have tried to come up with ways of explaining the observed results without invoking a non-thermal effect:
\begin{itemize}
	\item The use of microwave absorption spectroscopy techniques that have previously shown resonance-like artifacts\cite{Resonances1987}. An effective charge of $q=10^7 e^-$ was found, which is very large for a $10^9$ atom structure; genome and protein charge can hardly explain $10^5 e^-$, which would be reduced further by ionic screening. The potential energy of such a sphere of charge exceeds the metabolic burden to build a virus by a factor of 10.
	\item Infrared thermography only determines the surface temperature. Based on heating patterns seen in other experiments\cite{Effects1950}, the observed 7 C temperature rise does not seem to conclusively rule out a thermal effect. 
	\item A lack of free-space dosimetry or simulation of power density means that the frequency-selective inactivation could have been due to the frequency response of the power amplifier, antenna gain, and absorption of the cuvette.
\end{itemize}

A few well-established non-resonant non-thermal effects have already been put to practical use. Electroporation is a common laboratory procedure where an enormous electric field causes ions in solution to diffuse across a host cell's membrane, reversibly rendering it permeable (among other, more subtle mechanisms; the minutia of poration are extraordinarily complex and beyond the author's understanding\cite{Theoretical2007}). Irreversible electroporation\cite{Irreversible2013} has recently been used in the clinic\cite{Nonthermal2013} for tumor ablation.% In vitro there is evidence to suggest that sub-cellular effects take place under truly tremendous fields.

The genome charge is almost entirely neutralized by the nucleocapsid. From an energetics perspective, \footnote{we can obtain the spring constant $k = \omega_{res}^2 m^*$ from the resonant frequency and therefore from the speed of sound and $m^*$, the reduced mass.} With relatively sane choices of charge screening and resonant frequency, the energy absorbed by the whole virus per cycle at 5 MV/m might vary from $10^{-5}$ to 100 times the thermal energy $k_b T$.

Compare to $~5 k_b T$ per protein for 1 capsid segment binding energy \cite{Energies2012} \cite{Weak2002} and 6 to 30 $k_b T$ to form a single pore in a lipid bilayer \cite{Atomistic2014}.
%17 or 78 kJ/mol / (boltzmannConstant * 310 K) / avogadroConstant = 6.5 to 30.3 kT

On the other hand, such a sub-angstrom movement only produces approx. 0.05\% deflection of the capsid, whereas a general value for host cells is about 3\% deflection. In addition, with a relaxation time on the order of 500 ps, it is difficult to imagine that the energy will be sufficiently concentrated to be destructive, as after the relaxation time the energy is a distant memory; only the mechanical effects that were induced during that time can remain.

@@

@@

In terms of host safety, Pakhomov \cite{Comparative} find no non-thermal effects on functioning frog heart pacemaker cells at 0.9 MV/m (although the field inside the tissue was not measured and could have been lower). Excellent recent \textit{in vivo} data by de Seze following continuous exposure to nanosecond pulse trains at an external field up to 3 MV/m \cite{Repeated2020} notes little acute effects, but a massive rate of long-term carcinogenicity. Reviews \cite{Penetration2016c} \cite{Effects2016}.

@@

We hoped to shape the nanosecond pulse into a high-voltage 10 GHz tone to drive the virus. This requires an edge faster than ~100 ps. It is difficult, but not impossible\cite{Fundamental1998} to make spark gaps produce such fast edges without high-pressure\cite{Design2007d}\cite{Picosecond1993} or building the gap into a Marx\cite{Simple1991} pulse-forming line.

Avalanche transistors appear to be the easiest way to produce high power sub-100 picosecond pulses. We stole an avalanche Marx design verbatim from Li (\cite{Development2016b} \cite{Design2018c}), which allows large pulses from safer lower voltage supplies (albeit requiring somewhat costly capacitors); but

@@

We are particularly excited by the prospect of using picosecond time-domain measurements not only to perform steady state dielectric measurements or monitor the input pulse, as is now common practice, but to resolve the internal effects of a single high-voltage pulse on the virus in real time. Scan-converter based transient digitizers, such as the Tektronix SCD-5000 series, appear to be the lowest-cost method of obtaining such THz sampling bandwidths.

%When high voltage pulses are available, miniaturized \ntilde 2 mm internal diameter continuously pumped scan converter tubes like Fecher\cite{Production1949} micro-oscillographs using in-vacuum CMOS image sensors for direct-electron-detection appear to be an easy way to digitize 20+ GHz transients. The main bandwidth limitation appears to be the length of the Y deflection plate and the sensitivity of the recording medium.
